### 2022-12-09 introduction/introduction.tex
Made outline into comments

======= OUTLINE =======

BACKGROUND
Survery of historical developments that inform dissertation
Write at a high school or early college biology level

I. Proteins are the molecular machines that drive biology
    A. Four major macromolecules and briefly discuss broad roles of each
        1. Nucleic acids
        2. Carbohydrates
        3. Lipids
        4. Protein
    B. Some major examples of proteins
        1. Hemoglobin and sickle cell
        2. Color blindness and opsins
        3. Lactase and lactose intolerance
II. Proteins are linear chains of amino acids with a three-dimensional structure
    A. Background on biochemistry, primary, secondary, and tertiary structure
    B. Structure-function relationships returning to previous hemoglobin example
    C. Structural biology seeks to determine protein structures to understand function
    D. Overview of structural methods including purification steps
        1. X-ray crystallography
        2. NMR
        3. cryo-EM
III. Disordered proteins do not adopt a stable structure
    A. Definition and discussion of protein disorder
        1. Structural heterogeneity, conformational ensembles, floppy
        2. Major sequence characteristics and structural "modes"?
    B. Disordered proteins are challenging to study with traditional methods
    C. Genome sequencing and computational tools revealed disordered proteins are ubiquitous
IV. Disordered proteins perform varied functions
    A. Flexibility confers several advantages
    A. Survey of disordered proteins and their functions
        1. Linkers
        2. Signaling
V. Disordered proteins are enriched in transcriptional machinery
    A. Introduce transcription via the central dogma
    B. Describe transcription factors, cofactors, transcriptional machinery
        1. Examples of different transcription factors
    C. Describe properties of transcription factors
        1. DNA binding and activation domains
        2. Different flavors of activation domains
    D. How transcription factors find their targets to initiate specific genetic programs is one of the major outstanding questions in molecular biology
VI. Disordered proteins are associated with phase separation
    A. Describe phase separation in general
    B. Describe phase separation as related to proteins
    C. Describe functions of phase separation and relationship to subcellular structures
VII. Phase separation via IDRs is a possible mechanism for initiating transcription
    A. Summarize Stadler's papers and others

AIMS
Discuss more recent developments in field and propose driving questions

I. Structure-function relationships of disordered proteins are unclear
    A. Examples of studies that have found specific cases of conservation in IDRs
    B. Alan Moses paper as reference in yeast
II. Use evolutionary analyses to identify conserved properties of disordered regions
    A. Conservation is powerful signal of function
        1. Examples of other studies that applied evolution
    B. Outline overall approach and make correspondence to chapters
        1. Identification of orthologous groups
        2. Application of evolutionary analyses

======= DRAFT =======

>>>>

%======= OUTLINE =======
%
%BACKGROUND
%Survery of historical developments that inform dissertation
%Write at a high school or early college biology level
%
%I. Proteins are the molecular machines that drive biology
%    A. Four major macromolecules and briefly discuss broad roles of each
%        1. Nucleic acids
%        2. Carbohydrates
%        3. Lipids
%        4. Protein
%    B. Some major examples of proteins
%        1. Hemoglobin and sickle cell
%        2. Color blindness and opsins
%        3. Lactase and lactose intolerance
%II. Proteins are linear chains of amino acids with a three-dimensional structure
%    A. Background on biochemistry, primary, secondary, and tertiary structure
%    B. Structure-function relationships returning to previous hemoglobin example
%    C. Structural biology seeks to determine protein structures to understand function
%    D. Overview of structural methods including purification steps
%        1. X-ray crystallography
%        2. NMR
%        3. cryo-EM
%III. Disordered proteins do not adopt a stable structure
%    A. Definition and discussion of protein disorder
%        1. Structural heterogeneity, conformational ensembles, floppy
%        2. Major sequence characteristics and structural "modes"?
%    B. Disordered proteins are challenging to study with traditional methods
%    C. Genome sequencing and computational tools revealed disordered proteins are ubiquitous
%IV. Disordered proteins perform varied functions
%    A. Flexibility confers several advantages
%    A. Survey of disordered proteins and their functions
%        1. Linkers
%        2. Signaling
%V. Disordered proteins are enriched in transcriptional machinery
%    A. Introduce transcription via the central dogma
%    B. Describe transcription factors, cofactors, transcriptional machinery
%        1. Examples of different transcription factors
%    C. Describe properties of transcription factors
%        1. DNA binding and activation domains
%        2. Different flavors of activation domains
%    D. How transcription factors find their targets to initiate specific genetic programs is one of the major outstanding questions in molecular biology
%VI. Disordered proteins are associated with phase separation
%    A. Describe phase separation in general
%    B. Describe phase separation as related to proteins
%    C. Describe functions of phase separation and relationship to subcellular structures
%VII. Phase separation via IDRs is a possible mechanism for initiating transcription
%    A. Summarize Stadler's papers and others
%
%AIMS
%Discuss more recent developments in field and propose driving questions
%
%I. Structure-function relationships of disordered proteins are unclear
%    A. Examples of studies that have found specific cases of conservation in IDRs
%    B. Alan Moses paper as reference in yeast
%II. Use evolutionary analyses to identify conserved properties of disordered regions
%    A. Conservation is powerful signal of function
%        1. Examples of other studies that applied evolution
%    B. Outline overall approach and make correspondence to chapters
%        1. Identification of orthologous groups
%        2. Application of evolutionary analyses
%
%======= DRAFT =======
